---
title: "Practical Machine Learning Assignment"
output: html_document
---

### Executive Summary

This assignment takes the wearable computing data from the *Weight Lifting Exercise* dataset, and using machine learning algorithms, it predicts the manner in which the user did the exercise.

```{r echo=FALSE}
setwd("~/Documents/Courses/Coursera/machine_learning/Assignment")
```

### Getting and cleaning data

First, the training and testing data are read into the file from the given links. From the datasets, we could observe that it contrains "NA", "#DIV/0", and empty spaces for missing data. So, these are specified along with the 'read' input command.

```{r}
training <- read.csv('pml-training.csv', na.strings=c("NA","#DIV/0!",""))
testing <- read.csv('pml-testing.csv', na.strings=c("NA","#DIV/0!",""))
```

It is observed that the training dataset consists of a lot of missing variables. In order to avoid them being present in the model, a guideline is developed to clean the dataset. If the number of missing values in a column is greater than 10,000 (among 19,622 rows of training dataset), then that particular column is removed from the dataset.

```{r}
library(fields)
library(caret)
library(randomForest)
library(rpart)
```

```{r}
temp <- t(stats(training))
missingcol<-apply(temp, 1, function(currow){all(is.na(currow)) || (currow["missing values"] > 10000)})
trainingclean <- training[!missingcol]
```

```{r echo=FALSE}
trainingclean <- cbind(trainingclean, training[,160])
b <- colnames(trainingclean)
colnames(trainingclean) <- c(b[1:56], "classe")
```

### Machine Learning and Prediction

A set is seed first for consistency. First, the results are predicted using the 'rpart' method.

```{r}
set.seed(123)
trainingclean <- trainingclean[c(-1)]
model1 <- train(classe~., method="rpart", data=trainingclean)
predictions <- predict(model1, newdata=trainingclean)
confusionMatrix(predictions, training$classe)
```

The accuracy of using the 'rpart' method is very low, about **49.56%**. That means, about 50% of the time, the prediction outcome is wrong if we used this approach. Hence this method is disgarded.

Next, 'Random Forest' algorithm is used on the training set to predict the outcome.

```{r}
model2 <- randomForest(classe~., data=trainingclean)
print(model2)

predictions <- predict(model2, newdata=trainingclean)
confusionMatrix(predictions, training$classe)
```

In this case, the accuracy of using the method in the sample is **100%**. And the OOB (out-of-bag or out-of-sample) error is also very low (about **0.06%**), that means, even if this model is used to predict a different dataset, the outcome will be **94%** accurate.

Hence the second model is considered to be a better fit to predict the outcome of the wearable computing data. This model is used on the test data and the outcomes are entered in the submission assignment.

